---
title: "Final Report"
subtitle: "Experiments and Causality"
author: "Colby Carter, Abhishek Agarwal, Tiffany Jaya"
date: \today
output: pdf_document
---

## Load the libraries

```{r, message=FALSE}
library(data.table) # fread
library(dplyr)
library(geonames)
library(lmtest) # coeftest
library(lubridate) # time conversion
library(sandwich) # vcovHC
library(RJSONIO)
```

## Helper functions

```{r}
convert_fctr_to_boolean <- function(col) {
  return(as.numeric(as.logical(col)))
}

convert_fctr_to_datetime <- function(col) {
  return(as.POSIXct(col, format="%Y-%m-%d %H:%M:%S"))
}

convert_fctr_to_numeric <- function(col) {
  return(as.numeric(levels(col)[col]))
}

convert_fctr_to_str <- function(col) {
  return(as.character(col))
}
```

## Load the dataset

```{r}
# pilot study
d <- read.csv(file = "./W241 Colby Carter, Tiffany Jaya, Abhishek Agarwal_April 5, 2018_00.27.csv",
              header = TRUE, 
              sep = ",")
```

```{r}
# actual study
d <- read.csv(file = "./W241 Colby Carter, Tiffany Jaya, Abhishek Agarwal_April 12, 2018_09.41.csv",
              header = TRUE, 
              sep = ",")
```

## Clean up the dataset

**WARNING: DO NOT RERUN THIS SECTION TWICE!**

```{r}
# rename columns to be more descriptive
setnames(d, 
         old = c("Q1.1", 
                 "Q2.2_1",
                 "Q2.3_1",
                 "Q2.5_1",
                 "Q2.7_1",
                 "Q3.1_1", 
                 "Q3.3_1",
                 "Q3.5_1",
                 "Q3.7_1",
                 "Q4.1",
                 "Q4.2",
                 "Q4.3",
                 "Q4.4",
                 "Q4.5",
                 "Q4.6",
                 "Q4.7",
                 "Q4.8",
                 "Q4.9",
                 "Q4.10",
                 "Q4.11",
                 "Q4.12",
                 "Q4.13",
                 "Q4.14",
                 "Q4.15",
                 "Q4.16", 
                 "Q2.1_First.Click",
                 "Q2.1_Last.Click",
                 "Q2.1_Page.Submit",
                 "Q2.1_Click.Count",
                 "Q2.4_First.Click",
                 "Q2.4_Last.Click",
                 "Q2.4_Page.Submit",
                 "Q2.4_Click.Count",
                 "Q2.6_First.Click",
                 "Q2.6_Last.Click",
                 "Q2.6_Page.Submit",
                 "Q2.6_Click.Count",
                 "Q2.8_First.Click",
                 "Q2.8_Last.Click",
                 "Q2.8_Page.Submit",
                 "Q2.8_Click.Count",
                 "Q3.2_First.Click",
                 "Q3.2_Last.Click",
                 "Q3.2_Page.Submit",
                 "Q3.2_Click.Count",
                 "Q3.4_First.Click",
                 "Q3.4_Last.Click",
                 "Q3.4_Page.Submit",
                 "Q3.4_Click.Count",
                 "Q3.6_First.Click",
                 "Q3.6_Last.Click",
                 "Q3.6_Page.Submit",
                 "Q3.6_Click.Count",
                 "Q3.8_First.Click",
                 "Q3.8_Last.Click",
                 "Q3.8_Page.Submit",
                 "Q3.8_Click.Count"),
         new = c("consent", 
                 "control_employment",
                 "control_education", # baseline
                 "control_retirement",
                 "control_cybersecurity",
                 "treatment_employment",
                 "treatment_education", # baseline,
                 "treatment_retirement",
                 "treatment_cybersecurity",
                 "gender",
                 "age",
                 "highest_education",
                 "employment_status",
                 "marital_status",
                 "zip_code",
                 "community", # rural, urban, suburban
                 "can_vote",
                 "political_party",
                 "ethnicity",
                 "income",
                 "have_kids",
                 "internet_from_mobile",
                 "internet_from_home",
                 "internet_from_work",
                 "who_pays_internet",
                 "control_employment_First.Click",
                 "control_employment_Last.Click",
                 "control_employment_Page.Submit",
                 "control_employment_Click.Count",
                 "control_education_First.Click",
                 "control_education_Last.Click",
                 "control_education_Page.Submit",
                 "control_education_Click.Count",
                 "control_retirement_First.Click",
                 "control_retirement_Last.Click",
                 "control_retirement_Page.Submit",
                 "control_retirement_Click.Count",
                 "control_cybersecurity_First.Click",
                 "control_cybersecurity_Last.Click",
                 "control_cybersecurity_Page.Submit",
                 "control_cybersecurity_Click.Count",
                 "treatment_employment_First.Click",
                 "treatment_employment_Last.Click",
                 "treatment_employment_Page.Submit",
                 "treatment_employment_Click.Count",
                 "treatment_education_First.Click",
                 "treatment_education_Last.Click",
                 "treatment_education_Page.Submit",
                 "treatment_education_Click.Count",
                 "treatment_retirement_First.Click",
                 "treatment_retirement_Last.Click",
                 "treatment_retirement_Page.Submit",
                 "treatment_retirement_Click.Count",
                 "treatment_cybersecurity_First.Click",
                 "treatment_cybersecurity_Last.Click",
                 "treatment_cybersecurity_Page.Submit",
                 "treatment_cybersecurity_Click.Count"))
```

```{r}
# 1. remove the first two rows
d <- tail(d, -2)
# 2. remove rownames to avoid confusion (because it is not the subject's id)
rownames(d) <- NULL
# 3. safely convert columns of type factors to their respective types
# 3a. factor -> datetime
cols <- c("StartDate", "EndDate", "RecordedDate")
d[,cols] <- lapply(d[,cols], convert_fctr_to_datetime)
# 3b. factor -> logical/boolean
cols <- c("Finished")
d[,cols] <- lapply(d[,cols], convert_fctr_to_boolean)
d$consent <- ifelse(d$consent == "Yes", 1, 0)
d$can_vote <- ifelse(d$can_vote == "Yes", 1, 0)
d$have_kids <- ifelse(d$have_kids == "Yes", 1, 0)
# 3c. factor -> numeric
cols <- c("Progress", 
          "Duration..in.seconds.", 
          "LocationLatitude", 
          "LocationLongitude",
          "control_employment", 
          "control_education", 
          "control_retirement", 
          "control_cybersecurity",
          "treatment_employment", 
          "treatment_education", 
          "treatment_retirement", 
          "treatment_cybersecurity",
          "zip_code",
          "control_employment_First.Click", 
          "control_employment_Last.Click", 
          "control_employment_Page.Submit", 
          "control_employment_Click.Count", 
          "control_education_First.Click",
          "control_education_Last.Click",
          "control_education_Page.Submit",
          "control_education_Click.Count",
          "control_retirement_First.Click",
          "control_retirement_Last.Click",
          "control_retirement_Page.Submit",
          "control_retirement_Click.Count",
          "control_cybersecurity_First.Click",
          "control_cybersecurity_Last.Click",
          "control_cybersecurity_Page.Submit",
          "control_cybersecurity_Click.Count",
          "treatment_employment_First.Click",
          "treatment_employment_Last.Click",
          "treatment_employment_Page.Submit",
          "treatment_employment_Click.Count",
          "treatment_education_First.Click",
          "treatment_education_Last.Click",
          "treatment_education_Page.Submit",
          "treatment_education_Click.Count",
          "treatment_retirement_First.Click",
          "treatment_retirement_Last.Click",
          "treatment_retirement_Page.Submit",
          "treatment_retirement_Click.Count",
          "treatment_cybersecurity_First.Click",
          "treatment_cybersecurity_Last.Click",
          "treatment_cybersecurity_Page.Submit",
          "treatment_cybersecurity_Click.Count"
          )
d[,cols] <- lapply(d[,cols], convert_fctr_to_numeric)
# 3d. factor -> str
cols <- c("IPAddress", "ResponseId", "UserLanguage")
d[,cols] <- lapply(d[,cols], convert_fctr_to_str)
```

## Add additional information into the dataset

```{r, echo=FALSE}
# login information
options(geonamesUsername="sullivannicole")
```

```{r}
# get country information based on latlong
country_names <- mapply(
  function(lat, long) {
    suppressWarnings(try(GNcountryCode(lat, long)$countryName, silent = TRUE))
  }, d$LocationLatitude, d$LocationLongitude)
d$country <- country_names
```

## List all the columns

```{r}
colnames(d)
```

## Analysis

data: contains the original dataset
```{r}
# save a copy of the original dataset
data <- d
```

### Filter

d: contains the dataset after we filtered out for the following:

1. subjects who did not finish the survey 100%
2. subjects who consented to take the survey

```{r}
# filter out subjects who consented to take the survey
d <- d[which(d$consent == 1),]
```

3. subjects who are from USA

```{r}
# filter out subjects who are from USA
d <- d %>% filter(., country == "United States")
```

4. subjects who took the survey once
  * if subjects took the survey more than once, we only look at their first time taking the survey

```{r}

```

### How many took the survey?

```{r}
# how many subjects take the survey?
nrow(data)
```

638 subjects take the survey.

### Verify that ResponseId is unique

```{r}
# is ResponseId unique? 
# if yes, use ResponseId as a unique id
freq_responseid <- data %>% group_by(ResponseId) %>% summarize(freq = sum(n()))
sum(freq_responseid$freq > 1) == 0
```

ResponseId is unique. 

### Verify that all the subjects completed the survey

```{r}
# how many subjects did not finish the survey?
sum(!data$Finished)
```

```{r out.width="49%", fig.align="center", fig.show="hold"}
# plot the progress of each subject
hist(data$Progress, 
     xlab = "progress (%)",
     ylab = "number of users",
     breaks = -0.5:100.5, 
     ylim = c(0, 700),
     labels = TRUE)
```

All of the subjects completed the survey. 

### Verify that all subjects completed the survey in reasonable time (greater than 2 minutes)

```{r}
# how long does it take the subjects who consented to finish the survey in minutes?
subjects.consent <- data[which(data$consent == 1),]
summary(subjects.consent$Duration..in.seconds./60)
```

The following checks are unnecessary:

```{r}
# list all the subjects who take less than 2 minutes to finish the survey
subjects.less_2_min <- data[which(data$Duration..in.seconds./60 < 2),]
nrow(subjects.less_2_min)
```

```{r}
# list all subjects who did not consent
subjects.no_consent <- data[which(data$consent == 0),]
nrow(subjects.no_consent)
```

```{r}
# verify that the subjects who did not consent 
# match the subjects who took the survey in less than 2 minutes
all(subjects.less_2_min$ResponseId %in% subjects.no_consent$ResponseId)
```

### Verify that all subjects have English set as their language

```{r}
# list all languages not in English
data[which(data$UserLanguage != "EN"),]$UserLanguage
```

### Verify that all subjects are from United States

```{r}
# how many subjects are from United States?
data %>% filter(., country == "United States") %>% count(.)
# how many subjects are from outside the United States?
data %>% filter(., country != "United States") %>% count(.)
```

Out of 638 subjects, 580 subjects are from the United States and 58 are from outside the United States.

### Verify that subjects only took the survey once

```{r}
# 
```


```{r}
# check if subjects in control answer all questions
subjects.usa %>%
  select(control_employment, control_education, control_retirement, control_cybersecurity) %>%
  summarise_all(funs(sum(is.na(.))))
```

```{r}
# check if subjects in treatment answer all questions
subjects.usa %>%
  select(treatment_employment, treatment_education, treatment_retirement, treatment_cybersecurity) %>%
  summarise_all(funs(sum(is.na(.))))
```

# TODO 1

people with incomplete data
38.130.207.83
172.58.201.123
110.54.188.251
24.62.129.49
97.90.210.58
172.73.12.254
108.52.204.129
107.77.196.93
67.186.59.98

```{r}
subjects.usa %>% 
  group_by(IPAddress) %>% 
  summarize(freq = sum(n())) %>%
  filter(freq > 1)
```


```{r}
subjects.usa %>%
  filter(IPAddress == "38.130.207.83") %>%
  select(control_employment, control_education, control_retirement, control_cybersecurity)
```


```{r}
# filter out only people with complete data
```

# TODO 2

calculate length of time page_submit - first_click
can use click.count as covariate
can use length of time as covariate





```{r}
# filter out the treatment

# filter out the control
```



# TODO 3

1) a pre-analysis/EDA, just looking at some plots of scores for each answer, with a few cuts by covariates (e.g. histograms by political preference, gender; box-whisker plots for age groups, etc.)

# TODO 4

2) then, for each question, starting with our education placebo test, we have a _stargazer_ output that has 3 regressions: a) with only the treatment dummy, b) with treatment plus key covariates, and c) with treatment, key covariates AND interaction terms (i.e. HTEs)


## References

1. renaming column names: http://rprogramming.net/rename-columns-in-r/ 

## Unused stuff

### IP
Get state, country information from IPs. IPs are not reliable though. Plus, it's dang slow.

```
freegeoip <- function(ip, format = ifelse(length(ip)==1,'list','dataframe'))
{
    if (1 == length(ip))
    {
        # a single IP address
        url <- paste(c("http://api.ipstack.com/", ip, "?access_key=0e93a4defcd645bd8829ad75a7223437&output=json&legacy=1/"), collapse='')
        ret <- fromJSON(readLines(url, warn=FALSE))
        if (format == 'dataframe')
            ret <- data.frame(t(unlist(ret)))
        return(ret)
    } else {
        ret <- data.frame()
        for (i in 1:length(ip))
        {
            r <- freegeoip(ip[i], format="dataframe")
            ret <- bind_rows(ret, r)
        }
        return(ret)
    }
}   

try.ip <- function(ip) {
  suppressWarnings(try(freegeoip(ip), silent = TRUE))
}

ips <- try.ip(data$IPAddress)
```

### LatLong

Get state, country information from Google API. Unfortunately, it's taking too long and Google has a limit for the number of API calls.

How to call the API via the browser:
https://maps.googleapis.com/maps/api/geocode/json?latlng=40.714224,-73.961452&key=AIzaSyC8Q7FY6SzmEfpMTOcx2BTrnYaRkzX12Ik

How to call the API via R:
```
library(revgeo)
revgeo(longitude=-73.961452, latitude=40.714224, output="frame", provider="google", API="AIzaSyC8Q7FY6SzmEfpMTOcx2BTrnYaRkzX12Ik")
```

This will always fail because of the limitation to the number of API calls:
```
ll <- revgeo(longitude = data$LocationLongitude, latitude = data$LocationLatitude, output = "frame", provider = "google", API = "AIzaSyC8Q7FY6SzmEfpMTOcx2BTrnYaRkzX12Ik")
```

Getting USA (or any other country)'s latlong

```
library(ggmap)
geocode("USA")
```
