---
title: "Final Report"
subtitle: "Experiments and Causality"
author: "Colby Carter, Abhishek Agarwal, Tiffany Jaya"
date: \today
output: pdf_document
---

# Pilot study 1

```{r}
d <- read.csv(file="./W241 Colby Carter, Tiffany Jaya, Abhishek Agarwal_April 5, 2018_00.27.csv",
              header=TRUE, 
              sep=",")
```

```{r}
colnames(d)
```

* StartDate in America/Denver timezone informs us when the subject starts the survey.
* EndDate in America/Denver timezone informs us when the subject finishes the survey.
* IPAddress is the IPAddress of the subject.
* Progress in percentage is how far along subject took the survey. 
  We prefer if subject took 100% of the survey.
* Duration in seconds informs us how long the subject takes the survey.
* Finished in boolean is similar to Progress in that it informs us whether or not the subject finishes the survey.
* LocationLatitude and LocationLongitude inform us the latitude and longitude of the subject.
* UserLanguage will inform us what is the default subject's language is.
